Mega-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-09-24 paper/1_Initial/manuscript.qmd
fixed some typos

<!-- Summary -->
Study 1 establishes that once probelamtic items are removed the IAS is best conceptualized as a set of item clusters rather than a unidimensional scale. This structural insight provides the foundation for Study 2, where dispositional correlates are examined to further evaluate the IAS’s construct validity.

# Study 2

@@

<!-- AIms -->
The present study aimed to provide a comphrensive evaluation of the IAS, spanning both its structure (study 1) and its dispositional correlates (study 2). Together, these studies offer novel insights into the psychometric characteristics, conceptual interpretation, and practical applications of the IAS within the broader landscape of interoception research.

@@

Study 2 reinforced this domain-specific view. Certain item pairings, particularly Hunger–Thirst and Bruise–Blood sugar, consistently predicted mood and interoceptive outcomes as well as, or better than, the full scale. These findings indicate that interoceptive self-report is most informative when tied to ecologically meaningful bodily signals, echoing prior suggestions that physiological context are important in the measurement of interoception[@vlemincx2023novel]. At the same time, other pairings (e.g., cough–sneeze, wind–burp) provided little predictive value. Taken together with Study 1, this suggests that the strength of interoceptive reporting lies not in broad aggregation but in the careful targeting of specific bodily domains.

@@

Taken together, these studies advance understanding of the IAS and interoceptive self-report more broadly. The refined 14-item IAS provides a more efficient and interpretable measure, capturing domain-specific interoceptive signals that are differentially associated with psychological outcomes. Consistent associations with mood, alongside modest links to personality and beliefs, indicate that the IAS reflects adaptive, context-sensitive aspects of bodily self-perception rather than broad dispositional traits. At the same time, current findings highlight limitations in modality coverage and item specificity, suggesting that future scale development should incorporate multimodal, context-sensitive, and continuous analogue ratings or wider discrete scales, to better capture the richness of interoceptive processes and clarify their role in affective, cognitive, and health-related outcomes [@desmedt2022measures; @vlemincx2021].

>>>>

<!-- Summary -->
Study 1 establishes that once probeleamtic items are removed the IAS is best conceptualized as a set of item clusters rather than a unidimensional scale. This structural insight provides the foundation for Study 2, where dispositional correlates are examined to further evaluate the IAS’s construct validity.
# Study 2

@@

<!-- AIms -->
The present study aimed to provide a comprehensive evaluation of the IAS, spanning both its structure (study 1) and its dispositional correlates (study 2). Together, these studies offer novel insights into the psychometric characteristics, conceptual interpretation, and practical applications of the IAS within the broader landscape of interoception research.

@@

Study 2 reinforced this domain-specific view. Certain item pairings, particularly Hunger–Thirst and Bruise–Blood sugar, consistently predicted mood and interoceptive outcomes as well as, or better than, the full scale. These findings indicate that interoceptive self-report is most informative when tied to ecologically meaningful bodily signals, echoing prior suggestions that physiological context are important in the measurement of interoception [@vlemincx2023novel]. At the same time, other pairings (e.g., cough–sneeze, wind–burp) provided little predictive value. Taken together with Study 1, this suggests that the strength of interoceptive reporting lies not in broad aggregation but in the careful targeting of specific bodily domains.

@@

Taken together, these studies advance understanding of the IAS and interoceptive self-report more broadly. The refined 14-item IAS provides a more efficient and interpretable measure, capturing domain-specific interoceptive signals that are differentially associated with psychological outcomes. Consistent associations with mood, alongside modest links to personality and beliefs, indicate that the IAS reflects adaptive, context-sensitive aspects of bodily self-perception rather than broad dispositional traits. At the same time, current findings highlight limitations in modality coverage and item specificity, suggesting that future scale development should incorporate multimodal, context-sensitive, and continuous analogue ratings or wider discrete scales, to better capture the richness of interoceptive processes and clarify their role in affective, cognitive, and health-related outcomes [@desmedt2022measures; @vlemincx2023novel].
